Project assignment - Tahoe-LAFS with SHA-3 candidates

### 2010-10-27 paper/main.tex
Added NIST evaluation criterias

\begin{acronym}
\acro{AES}{Advanced Encryption Standard}
\acro{LAFS}{Least-Authority Filesystem}
\acro{FEC}{Forward Error Correction}
\end{acronym}

@@

\section{Cryptographic Hash Functions}
A Cryptographic Hash function is a deterministic mathematical procedure which
takes an arbritary block of data and outputs a fixed-size bit string. The
output is refered to as the hash value, message digest or simply digest.
Another property of a cryptographic hash function is that a small change in the
input data (just 1 bit) should completly change the output of the hash
function. In other words it should be infeasible to find the reverse of a
cryptographic hash function. It should also be infeasible to find two blocks of
data which produce the same hash value (a collision).

\subsection{NIST SHA-3 Competition}
% Hash functions, NIST Comptetition, Candidates, SuperCop?
The Secure Hash Function version 3 (SHA-3) is a comming standard cryptographic
hash function set to replace the current standards SHA-1 and SHA-2.
It will be selected from a competition held by the National
Institute of Standards and Technology (NIST). The competition started in 2006
with submission deadline for candidates at the end of 2008. 51 candidates made
it into round 1, however just 14 passed to the second round. The reason for not
passing round 1 are either design weaknesses, substantial cryptographic
weaknesses or performance issues. The winner is sceduled to be selected by the 
2nd quarter of 2012.

The submissions for the various rounds of the SHA-3 competition include
reference implementations, optimized implementations, test-vectors to verify
implementations and documentations explaining the submission.

@@

\subsection{Architecture}

Tahoe has a three layer architecture: the key-value store, the filesystem, and
the application.

The \textbf{key-value store}, or the ``capability-data bytes'' store, is the
lowest layer and is implemented by a grid of Tahoe-LAFS storage servers. Data is
kept on the storage servers in the form of ``shares'', which are encrypted and
encoded parts of files. Capabilities are short ASCII strings, containing
information on where to \emph{find} a file, and how to \emph{verify} it.
Nodes in the grid learn about each other through an ``introducer'', which
roughly relates to a tracker in the BitTorrent\footnote{See
\url{http://www.bittorrent.org/beps/bep\_0003.html}} protocol.

The \textbf{filesystem} layer is responsible for mapping human-meaningful
pathnames to pieces of data. Each directory contains a table of capabilities
for its children, i.e. subdirectories or files. Two forms of capabilities is
available for each file, read-only and read-write, and these can be shared to
provide shared/published directory structures with friends.

Since it is not practical for users to remember strings containing random
characters, the \textbf{application} layer is used for providing a user-friendly
interface to the directories and files.

\subsection{Related technologies}

Tahoe-\ac{LAFS} uses a number of techniques to realize the functionality
demanded by a secure, distributed filesystem.

\paragraph{Erasure coding.}

When a client puts a file on the grid, it first encrypts the file, before
breaking the file into small segments. The segments are then erasure coded.
The use of \emph{erasure coding} enables Tahoe to recover a file using only a
predefined subset of the parts distributed to the storage servers, i.e. the
other nodes in the grid. Erasure coding is a type of \ac{FEC} code, which
extends a message with $C$ characters into a longer message with $N$ symbols
\cite{t_reed-solomon}.
The original $C$ characters can then be recovered from a subset of the $N$
symbols.

The properties of erasure coding can be seen up on as those of recovery in RAID
systems. \cite{t_erasure} % TODO: comparison with RAID

\paragraph{Merkle hash trees.}

Describe hash trees in further detail?

@@

\begin{figure}[h]
    \centering
    \includegraphics[width=0.9\columnwidth]{Tahoe-hashing.pdf}
    \caption{Tahoe-LAFS: Location of hashing}
    \label{fig:tahoehashing}
\end{figure}

>>>>

\begin{acronym}
\acro{AES}{Advanced Encryption Standard}
\acro{FEC}{Forward Error Correction}
\acro{LAFS}{Least-Authority Filesystem}
\acro{NIST}{National Institute of Standards and Technology}
\acro{RAM}{Random Access Memory}
\acro{SHA}{Secure Hash Algorithm}
\end{acronym}

@@

\section{Cryptographic Hash Functions}
% Get a source for this %
A Cryptographic Hash function is a deterministic mathematical procedure which
takes an arbritary block of data and outputs a fixed-size bit string. The
output is refered to as the hash value, message digest or simply digest.
Another property of a cryptographic hash function is that a small change in the
input data (just 1 bit) should completly change the output of the hash
function. In other words it should be infeasible to find the reverse of a
cryptographic hash function. It should also be infeasible to find two blocks of
data which produce the same hash value (a collision).

\subsection{NIST SHA-3 Competition}
The \ac{SHA} version 3 is a comming standard set to superceed the current
standards \ac{SHA}-1 and \ac{SHA}-2 families. The hash function to become
\ac{SHA}-3 will be decided by \ac{NIST} and choosen between the submitted
contestants to the \ac{NIST} hash competition. The current status of the 
competition is officially called Round 2, with 14 of 64 candidates having 
survived round 1\cite{s_fedreg}.

\subsubsection{\ac{NIST} evaluation criterias for \ac{SHA}-3}

\paragraph{Security} is the most important criteria for the SHA-3
candidates\cite{s_nistround2}. The full description of security criterias are
mentioned in \cite{s_fedreg}, however the most noteworthy criteria in regards
to Tahoe-\ac{LAFS} is that \ac{SHA}-3 candidates are required to have
resistance against length-extension attacks. Length-extension attacks is
something that both \ac{SHA}-1 and the \ac{SHA2}-family suffers
from\cite{s_mdre}.

\paragraph{Cost and Performance} are considered to be the 2nd most important
criteria. The absolute minimum for performance is that the SHA-3 candidate
should be faster than the functions in the \ac{SHA}-2 family. Cost is a measure
of how much memory an implementation requires in software, both the 
implementation itself and the use of \ac{RAM} during runtime. Another measure
of cost is how many logic gates it takes to implement the function in hardware.

\paragraph{Algorithm and implementation characteristics} is the third criteria.
By this criteria \ac{NIST} emphasis that algorithms with greater flexbility
will be given preference over other algorithms\cite{s_nistround2}. By flexible
they imply the possiblity of the function to run efficiently on a varity of
plattforms, to use parallelism or instruction set extensions. Another key point
of a flexible hash function is that it should have a simple and elegant design
to encourage understanding, analysis and design confidence.

@@

\subsection{Erasure coding}

The use of \emph{erasure coding} enables Tahoe to recover a file using only a
predefined subset of the parts distributed to the storage servers, i.e. the
other nodes in the grid. Erasure coding is a type of \ac{FEC} code, which
extends a message with $C$ characters into a longer message with $N$ symbols.
The original $C$ characters can then be recovered from a subset of the $N$
symbols.

The properties of erasure coding can be seen up on as those of recovery in RAID
systems.

% TODO: comparison with RAID?

@@

% Describe hash trees in further detail?

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.9\columnwidth]{Tahoe-hashing.pdf}
    \caption{Tahoe-LAFS: Location of hashing}
    \label{fig:tahoehashing}
\end{figure}
